---
title: "Titanic arules"
output: html_notebook
---

#Analiza univariata
##Data shape
```{r}
summary(data)
```

##Age
```{r}
# Mean
mean(data$Age, na.rm = TRUE)
```
```{r}
# Median
median(data$Age, na.rm = TRUE)
```
```{r}
x <- na.omit(data$Age)
h<-hist(x, breaks = 80, col = "grey", xlab="Age", main="Age Histogram")
xfit<-seq(min(x),max(x),length=40) 
yfit<-dnorm(xfit,mean=mean(x),sd=sd(x)) 
yfit <- yfit*diff(h$mids[1:2])*length(x) 
lines(xfit, yfit, col="red", lwd=3)
```

##Sex
```{r}
sexFrequency = table(data$Sex)
ylim <- c(0, 1.2*max(sexFrequency))
b = barplot(sexFrequency, main = "Sex frequency", ylim = ylim, xlab="Sex", ylab = "Frequency")
text(x = b, y = sexFrequency, label = sexFrequency, pos = 3, col = "blue")
```
##Ticket class
```{r}
classFrequency = table(data$Pclass)
names(classFrequency) = c("1st","2nd","3rd")
ylim <- c(0, 1.2*max(classFrequency))
b = barplot(classFrequency, main = "Ticket class frequency", ylim = ylim, xlab="Ticket class", ylab = "Frequency")
text(x = b, y = classFrequency, label = classFrequency, pos = 3, col = "blue")

```

##Fare
```{r}
# Mean
mean(data$Fare, na.rm = TRUE)
```
```{r}
# Median
median(data$Fare, na.rm = TRUE)
```
```{r}
x <- na.omit(data$Fare)
hist(x, breaks = 100, col = "grey", xlab="Fare ($)", main="Fare Histogram")
```

##Port of Embarkation
```{r}
classFrequency = table(data$Embarked)
names(classFrequency) = c('Uknown', 'Cherbourg', 'Queenstown', 'Southampton')
ylim <- c(0, 1.2*max(classFrequency))
b = barplot(classFrequency, main = "Port of Embarkation frequency", ylim = ylim, xlab="Ticket class", ylab = "Frequency")
text(x = b, y = classFrequency, label = classFrequency, pos = 3, col = "blue")
```

##Survival
```{r}
classFrequency = table(data$Survived)
names(classFrequency) = c("No","Yes")
ylim <- c(0, 1.2*max(classFrequency))
b = barplot(classFrequency, main = "Survival frequency", ylim = ylim, xlab="Survived", ylab = "Frequency")
text(x = b, y = classFrequency, label = classFrequency, pos = 3, col = "blue")
```

#Analiza bivariata

## Scatter plot fare vs age
```{r}
plot(data$Fare, data$Age, xlab = 'Fare', ylab = 'Age')
```
## Correlation test fare vs age
```{r}
cor.test(data$Age, data$Fare, method = 'pearson')
```
## Scatter plot fare vs class
```{r}
plot(data$Fare, data$Pclass, xlab = 'Fare', ylab = 'Age')
```

## Correlation test fare vs class
```{r}
cor.test(data$Fare, data$Pclass, method = 'pearson')
```

## Plot age vs sex
```{r}
ggplot(data, aes(Age, fill = Sex)) +
  geom_histogram(alpha = 0.5, aes(y = ..count..))

```
## Correlation test age vs sex
```{r}
 t.test(Age ~ Sex, data = data)
```

## Plot class vs sex
```{r}
ggplot(data, aes(Pclass, fill = Sex)) +
  geom_bar(alpha = 0.5, aes(y = ..count..))

```

## Correlation test class vs sex
```{r}
 t.test(Pclass ~ Sex, data = data)
```

## Mosaicplot survival vs class
```{r}
mosaicplot(
  data$Pclass ~ data$Survived,
  main = "Passenger Fate by Traveling Class",
  shade = FALSE,
  color = TRUE,
  xlab = "Passenger Class",
  ylab = "Survived"
)

```
## Correlation test survival vs class
```{r}
 t.test(Pclass ~ Survived, data = data)
```

## Correlation test survival vs class
```{r}
fisher.test(data$Survived, data$Pclass)
```

## Mosaicplot survival vs sex
```{r}
mosaicplot(
  data$Sex ~ data$Survived,
  main = "Passenger Fate by Sex",
  shade = FALSE,
  color = TRUE,
  xlab = "Sex",
  ylab = "Survived"
)

```
## Correlation test survival vs sex
```{r}
fisher.test(data$Survived, data$Sex)
```
## Age vs survival
```{r}
ggplot(data, aes(x=Age)) + geom_density(aes(color=Survived))
```

## Boxplot survived vs age
```{r}
boxplot(data$Age~data$Survived,main="Proportion of Survivors by Age", col=c('darkseagreen4','salmon4'), xlab="Survived", ylab="Age") 
```

## Mosaicplot survival vs port of embarked
```{r}
mosaicplot(
  data$Embarked ~ data$Survived,
  main = "Passenger Fate by Port of embarked",
  shade = FALSE,
  color = TRUE,
  xlab = "Port of embarked",
  ylab = "Survived"
)

```

## Correlation test survival vs port of embarked
```{r}
fisher.test(data$Survived, data$Embarked)
```

