---
title: "Analiza prin clasificare nesupervizata"
output: html_notebook
---


#Part 1
##2d-10c Dataset
##Best algorithm: hierarchic ward variance
##Load data
```{r}
library(dbscan)
library(cluster)
library(fpc)
library(factoextra)
library(NbClust)
library(mclust)
library(EMCluster, quietly = TRUE)

data <- as.vector(read.table("../tema4/2d-10c.dat", header=F, skip=3))
data <- data[-3]
print(data)
```
##Algoritmii ierarhici: single link
```{r}
d <- dist(data, method = "euclidean")
hc1 <- hclust(d, method = "single" )
# Plot the obtained dendrogram
plot(hc1, cex = 0.6, hang = -1)
rect.hclust(hc1, k = 10, border = 2:20)
clust <- cutree(hc1, k = 10)
fviz_cluster(list(data = data, cluster = clust))
```
##Algoritmii ierarhici: complete link
```{r}
d <- dist(data, method = "euclidean")
hc1 <- hclust(d, method = "complete" )
# Plot the obtained dendrogram
plot(hc1, cex = 0.6, hang = -1)
rect.hclust(hc1, k = 10, border = 2:20)
clust <- cutree(hc1, k = 10)
fviz_cluster(list(data = data, cluster = clust))
```
##Algoritmii ierarhici: average link
```{r}
d <- dist(data, method = "euclidean")
hc1 <- hclust(d, method = "average" )
# Plot the obtained dendrogram
plot(hc1, cex = 0.6, hang = -1)
rect.hclust(hc1, k = 10, border = 2:20)
clust <- cutree(hc1, k = 10)
fviz_cluster(list(data = data, cluster = clust))
```
##Algoritmii ierarhici: ward variance
```{r}
d <- dist(data, method = "euclidean")
hc1 <- hclust(d, method = "ward.D" )
# Plot the obtained dendrogram
plot(hc1, cex = 0.6, hang = -1)
rect.hclust(hc1, k = 10, border = 2:20)
clust <- cutree(hc1, k = 10)
fviz_cluster(list(data = data, cluster = clust))
```
##DBscan
```{r}
db = dbscan(data,0.5)
hullplot(data, db$cluster)
```
##K-means
```{r}
set.seed(7)
km1 = kmeans(data, 10 , nstart=20)
kmeans.res <- kmeans(data, centers = 10)
fviz_cluster(kmeans.res, data = data, geom = "point")
```
##EM
```{r}
set.seed(1234)

emobj <- simple.init(data, nclass = 10)
emobj <- shortemcluster(data, emobj)
summary(emobj)

ret <- emcluster(data, emobj, assign.class = TRUE)
summary(ret)
```

##Iris dataset
##Best algorithm: average linkage
##Load data
```{r}
library(dbscan)
library(cluster)
library(fpc)
library(factoextra)
library(NbClust)
library(mclust)

data <- read.csv(file="../tema4/iris.csv", header=F)
data <- data[-5]
print(data)
```
##Algoritmii ierarhici: single link
```{r}
d <- dist(data, method = "euclidean")
hc1 <- hclust(d, method = "single" )
# Plot the obtained dendrogram
plot(hc1, cex = 0.6, hang = -1)
rect.hclust(hc1, k = 10, border = 2:20)
clust <- cutree(hc1, k = 10)
fviz_cluster(list(data = data, cluster = clust))
```
##Algoritmii ierarhici: complete link
```{r}
d <- dist(data, method = "euclidean")
hc1 <- hclust(d, method = "complete" )
# Plot the obtained dendrogram
plot(hc1, cex = 0.6, hang = -1)
rect.hclust(hc1, k = 10, border = 2:20)
clust <- cutree(hc1, k = 10)
fviz_cluster(list(data = data, cluster = clust))
```
##Algoritmii ierarhici: average link
```{r}
d <- dist(data, method = "euclidean")
hc1 <- hclust(d, method = "average" )
# Plot the obtained dendrogram
plot(hc1, cex = 0.6, hang = -1)
rect.hclust(hc1, k = 10, border = 2:20)
clust <- cutree(hc1, k = 10)
fviz_cluster(list(data = data, cluster = clust))
```
##Algoritmii ierarhici: ward variance
```{r}
d <- dist(data, method = "euclidean")
hc1 <- hclust(d, method = "ward.D" )
# Plot the obtained dendrogram
plot(hc1, cex = 0.6, hang = -1)
rect.hclust(hc1, k = 10, border = 2:20)
clust <- cutree(hc1, k = 10)
fviz_cluster(list(data = data, cluster = clust))
```
##DBscan
```{r}
db = dbscan(data,0.5)
hullplot(data, db$cluster)
```
##K-means
```{r}
set.seed(7)
km1 = kmeans(data, 10 , nstart=20)
kmeans.res <- kmeans(data, centers = 10)
fviz_cluster(kmeans.res, data = data, geom = "point")
```
##EM
```{r}
set.seed(1234)

emobj <- simple.init(data, nclass = 10)
emobj <- shortemcluster(data, emobj)
summary(emobj)

ret <- emcluster(data, emobj, assign.class = TRUE)
summary(ret)
```

##Long dataset
##Best algorithm:DBScan
##Load data
```{r}
library(dbscan)
library(cluster)
library(fpc)
library(factoextra)
library(NbClust)
library(mclust)

data <- as.vector(read.table("../tema4/long.data", header=F))
data <- data[-3]
print(data)
```
##Algoritmii ierarhici: single link
```{r}
d <- dist(data, method = "euclidean")
hc1 <- hclust(d, method = "single" )
# Plot the obtained dendrogram
plot(hc1, cex = 0.6, hang = -1)
rect.hclust(hc1, k = 10, border = 2:20)
clust <- cutree(hc1, k = 2)
fviz_cluster(list(data = data, cluster = clust))
```
##Algoritmii ierarhici: complete link
```{r}
d <- dist(data, method = "euclidean")
hc1 <- hclust(d, method = "complete" )
# Plot the obtained dendrogram
plot(hc1, cex = 0.6, hang = -1)
rect.hclust(hc1, k = 2, border = 2:20)
clust <- cutree(hc1, k = 2)
fviz_cluster(list(data = data, cluster = clust))
```
##Algoritmii ierarhici: average link
```{r}
d <- dist(data, method = "euclidean")
hc1 <- hclust(d, method = "average" )
# Plot the obtained dendrogram
plot(hc1, cex = 0.6, hang = -1)
rect.hclust(hc1, k = 2, border = 2:20)
clust <- cutree(hc1, k = 2)
fviz_cluster(list(data = data, cluster = clust))
```
##Algoritmii ierarhici: ward variance
```{r}
d <- dist(data, method = "euclidean")
hc1 <- hclust(d, method = "ward.D" )
# Plot the obtained dendrogram
plot(hc1, cex = 0.6, hang = -1)
rect.hclust(hc1, k = 2, border = 2:20)
clust <- cutree(hc1, k = 2)
fviz_cluster(list(data = data, cluster = clust))
```
##DBscan
```{r}
db = dbscan(data,0.2)
hullplot(data, db$cluster)
```
##K-means
```{r}
set.seed(7)
km1 = kmeans(data, 2 , nstart=20)
kmeans.res <- kmeans(data, centers = 2)
fviz_cluster(kmeans.res, data = data, geom = "point")
```
##EM
```{r}
set.seed(1234)

emobj <- simple.init(data, nclass = 2)
emobj <- shortemcluster(data, emobj)
summary(emobj)

ret <- emcluster(data, emobj, assign.class = TRUE)
summary(ret)
```

##order2-3clust dataset
##Load data
```{r}
library(dbscan)
library(cluster)
library(fpc)
library(factoextra)
library(NbClust)
library(mclust)

data <- read.csv(file="../tema4/order2-3clust.csv", header=F)
data <- data[-3]
print(data)
```
##Algoritmii ierarhici: single link
```{r}
d <- dist(data, method = "euclidean")
hc1 <- hclust(d, method = "single" )
# Plot the obtained dendrogram
plot(hc1, cex = 0.6, hang = -1)
rect.hclust(hc1, k = 10, border = 2:20)
clust <- cutree(hc1, k = 3)
fviz_cluster(list(data = data, cluster = clust))
```
##Algoritmii ierarhici: complete link
```{r}
d <- dist(data, method = "euclidean")
hc1 <- hclust(d, method = "complete" )
# Plot the obtained dendrogram
plot(hc1, cex = 0.6, hang = -1)
rect.hclust(hc1, k = 3, border = 2:20)
clust <- cutree(hc1, k = 3)
fviz_cluster(list(data = data, cluster = clust))
```
##Algoritmii ierarhici: average link
```{r}
d <- dist(data, method = "euclidean")
hc1 <- hclust(d, method = "average" )
# Plot the obtained dendrogram
plot(hc1, cex = 0.6, hang = -1)
rect.hclust(hc1, k = 3, border = 2:20)
clust <- cutree(hc1, k = 3)
fviz_cluster(list(data = data, cluster = clust))
```
##Algoritmii ierarhici: ward variance
```{r}
d <- dist(data, method = "euclidean")
hc1 <- hclust(d, method = "ward.D" )
# Plot the obtained dendrogram
plot(hc1, cex = 0.6, hang = -1)
rect.hclust(hc1, k = 3, border = 2:20)
clust <- cutree(hc1, k = 3)
fviz_cluster(list(data = data, cluster = clust))
```
##DBscan
```{r}
db = dbscan(data,0.2)
hullplot(data, db$cluster)
```
##K-means
```{r}
set.seed(7)
km1 = kmeans(data, 3 , nstart=20)
kmeans.res <- kmeans(data, centers = 3)
fviz_cluster(kmeans.res, data = data, geom = "point")
```
##EM
```{r}
set.seed(1234)

emobj <- simple.init(data, nclass = 3)
emobj <- shortemcluster(data, emobj)
summary(emobj)

ret <- emcluster(data, emobj, assign.class = TRUE)
summary(ret)
```
##smile dataset
##Best algorithm: Single linkage
##Load data
```{r}
library(dbscan)
library(cluster)
library(fpc)
library(factoextra)
library(NbClust)
library(mclust)

data <- read.csv(file="../tema4/smile.csv", header=F)
data <- data[-3]
print(data)
```
##Algoritmii ierarhici: single link
```{r}
d <- dist(data, method = "euclidean")
hc1 <- hclust(d, method = "single" )
# Plot the obtained dendrogram
plot(hc1, cex = 0.6, hang = -1)
rect.hclust(hc1, k = 4, border = 2:20)
clust <- cutree(hc1, k = 4)
fviz_cluster(list(data = data, cluster = clust))
```
##Algoritmii ierarhici: complete link
```{r}
d <- dist(data, method = "euclidean")
hc1 <- hclust(d, method = "complete" )
# Plot the obtained dendrogram
plot(hc1, cex = 0.6, hang = -1)
rect.hclust(hc1, k = 4, border = 2:20)
clust <- cutree(hc1, k = 4)
fviz_cluster(list(data = data, cluster = clust))
```
##Algoritmii ierarhici: average link
```{r}
d <- dist(data, method = "euclidean")
hc1 <- hclust(d, method = "average" )
# Plot the obtained dendrogram
plot(hc1, cex = 0.6, hang = -1)
rect.hclust(hc1, k = 4, border = 2:20)
clust <- cutree(hc1, k = 4)
fviz_cluster(list(data = data, cluster = clust))
```
##Algoritmii ierarhici: ward variance
```{r}
d <- dist(data, method = "euclidean")
hc1 <- hclust(d, method = "ward.D" )
# Plot the obtained dendrogram
plot(hc1, cex = 0.6, hang = -1)
rect.hclust(hc1, k = 4, border = 2:20)
clust <- cutree(hc1, k = 4)
fviz_cluster(list(data = data, cluster = clust))
```
##DBscan
```{r}
db = dbscan(data,0.2)
hullplot(data, db$cluster)
```
##K-means
```{r}
set.seed(7)
km1 = kmeans(data, 4 , nstart=20)
kmeans.res <- kmeans(data, centers = 4)
fviz_cluster(kmeans.res, data = data, geom = "point")
```
##EM
```{r}
set.seed(1234)

emobj <- simple.init(data, nclass = 4)
emobj <- shortemcluster(data, emobj)
summary(emobj)

ret <- emcluster(data, emobj, assign.class = TRUE)
summary(ret)
```

##square dataset
##Best algorithm: k-means
##Load data
```{r}
library(dbscan)
library(cluster)
library(fpc)
library(factoextra)
library(NbClust)
library(mclust)

data <- as.vector(read.table("../tema4/square.data", header=F))
data <- data[-3]
print(data)
```
##Algoritmii ierarhici: single link
```{r}
d <- dist(data, method = "euclidean")
hc1 <- hclust(d, method = "single" )
# Plot the obtained dendrogram
plot(hc1, cex = 0.6, hang = -1)
rect.hclust(hc1, k = 4, border = 2:20)
clust <- cutree(hc1, k = 4)
fviz_cluster(list(data = data, cluster = clust))
```
##Algoritmii ierarhici: complete link
```{r}
d <- dist(data, method = "euclidean")
hc1 <- hclust(d, method = "complete" )
# Plot the obtained dendrogram
plot(hc1, cex = 0.6, hang = -1)
rect.hclust(hc1, k = 4, border = 2:20)
clust <- cutree(hc1, k = 4)
fviz_cluster(list(data = data, cluster = clust))
```
##Algoritmii ierarhici: average link
```{r}
d <- dist(data, method = "euclidean")
hc1 <- hclust(d, method = "average" )
# Plot the obtained dendrogram
plot(hc1, cex = 0.6, hang = -1)
rect.hclust(hc1, k = 4, border = 2:20)
clust <- cutree(hc1, k = 4)
fviz_cluster(list(data = data, cluster = clust))
```
##Algoritmii ierarhici: ward variance
```{r}
d <- dist(data, method = "euclidean")
hc1 <- hclust(d, method = "ward.D" )
# Plot the obtained dendrogram
plot(hc1, cex = 0.6, hang = -1)
rect.hclust(hc1, k = 4, border = 2:20)
clust <- cutree(hc1, k = 4)
fviz_cluster(list(data = data, cluster = clust))
```
##DBscan
```{r}
db = dbscan(data,0.5)
hullplot(data, db$cluster)
```
##K-means
```{r}
set.seed(7)
km1 = kmeans(data, 4 , nstart=20)
kmeans.res <- kmeans(data, centers = 4)
fviz_cluster(kmeans.res, data = data, geom = "point")
```
##EM
```{r}
set.seed(1234)

emobj <- simple.init(data, nclass = 4)
emobj <- shortemcluster(data, emobj)
summary(emobj)

ret <- emcluster(data, emobj, assign.class = TRUE)
summary(ret)
```
#Part 2

##Load data
```{r}
library(cluster)
library(fpc)
library(factoextra)
library(NbClust)

data <- as.vector(read.table("../tema4/2d-10c.dat", header=F, skip=3))
data <- data[-3]
summary(data)
```

##k-Means
###Sum of squares
```{r}
k.max <- 20
wss <- sapply(1:k.max, function(k) {
    kmeans(data, centers = k, nstart = 20)$tot.withinss
  })
plot(1:k.max, wss,
     type="b", pch = 19, frame = FALSE, 
     xlab="Number of clusters K")
abline(v = 10, lty =2)

kmeans.res <- kmeans(data, centers = 10)
fviz_cluster(kmeans.res, data = data, geom = "point")
```

###Sillouette Width
```{r}
k.max <- 20
sil <- rep(0, k.max)
for(i in 2:k.max){
  km.res <- kmeans(data, centers = i, nstart = 20, iter.max=30)
  ss <- silhouette(km.res$cluster, dist(data))
  sil[i] <- mean(ss[, 3])
}

plot(1:k.max, sil, type = "b", pch = 19, 
     frame = FALSE, xlab = "Number of clusters k")
abline(v = which.max(sil), lty = 2)

kmeans.res <- kmeans(data, centers = which.max(sil))
fviz_cluster(kmeans.res, data = data, geom = "point")
```


##Hierarchical clustering
###Sum of squares
```{r}
dist.res <- dist(data, method = "euclidean")
hc <- hclust(dist.res, method = "complete")
k.max <- 20

wss <- function(d) {
  sum(scale(d, scale = FALSE)^2)
}

wrapper <- sapply(1:k.max, function(k) {
  cl <- cutree(hc, k)
  spl <- split(data, cl)
  wss <- sum(sapply(spl, wss))
  wss
})

plot(1:k.max, wrapper,
     type="b", pch = 19, frame = FALSE, 
     xlab="Number of clusters K")
abline(v = 5, lty =2)

plot(hc, labels = FALSE, hang = -1)
rect.hclust(hc, k = 5, border = 2:20)
```

###Sillouette Width
```{r}
dist.res <- dist(data, method = "euclidean")
hc <- hclust(dist.res, method = "complete")
k.max <- 20
sil <- rep(0, k.max)

for(i in 2:k.max){
  clusters <- cutree(hc, k = i)
  ss <- silhouette(clusters, dist(data))
  sil[i] <- mean(ss[, 3])
}

plot(1:k.max, sil, type = "b", pch = 19, 
     frame = FALSE, xlab = "Number of clusters k")
abline(v = which.max(sil), lty = 2)

plot(hc, labels = FALSE, hang = -1)
rect.hclust(hc, k = which.max(sil), border = 2:20)
```
